%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Submission for frontier optics.
% Authors:
% Xiaofeng Qian, Son H. Nguyen, Matthew Daddino, Ching Man Wong
% Stevens Institute of Technology
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% !TEX root = Thesis.tex

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
	a4paper,
	12pt,
	oneside,
]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[]{amsthm}
\usepackage[]{amssymb}
\usepackage[]{setspace}
\usepackage[left=1in, right=1in, top=0.6in, bottom=1in]{geometry}
\usepackage[colorlinks=true, linkcolor=blue, urlcolor=red, citecolor=blue]{hyperref}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{braket}
\usepackage{gensymb}
\usepackage{booktabs}

\onehalfspacing

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR
%----------------------------------------------------------------------------------------

\title{\textbf{Performing Quantum Computation with Classical Optics}}

\author{\small Son H. Nguyen\textsuperscript{1}, Matthew Daddino\textsuperscript{1}, Ching Man Wong\textsuperscript{1}, Xiaofeng Qian\textsuperscript{1}}

% remove printed date; keep a small affiliation line instead
\date{\small \textsuperscript{1}Physics Department, Stevens Institute of Technology, Hoboken, NJ 07030, USA}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle
\vspace{-1em}

%----------------------------------------------------------------------------------------
%	MANUSCRIPT (compact, experimental focus)
%----------------------------------------------------------------------------------------

\begin{abstract}
\noindent We present an experimental implementation of the Variational Quantum Eigensolver (VQE) using classical optical hardware. Qubits are encoded in photonic polarization and spatial modes; single-qubit rotations and mode couplings are realized with wave plates, beam splitters and dove prisms, and readout is performed via Stokes-parameter measurements. We demonstrate the optical construction of a UCCSD-inspired two-qubit ansatz for H$_2$, report state-preparation fidelities from Stokes-based tomography, and discuss stability and scalability trade-offs for photonic VQE prototypes.
\end{abstract}
This work translates a small-molecule VQE workflow into an experimentally accessible optical testbed. The goal is not to claim quantum advantage, but to provide a transparent platform for prototyping variational circuits, calibration strategies, and measurement protocols using standard optics.

We implement qubits using two degrees of freedom of a single photon: polarization (horizontal $\ket{H}=\ket{0}_p$, vertical $\ket{V}=\ket{1}_p$) and a two-mode spatial basis (labels $\ket{0}_s,\ket{1}_s$). The main components are:
- Quarter-wave plates (QWP) and half-wave plates (HWP) for polarization rotations (Jones-matrix control).
- Non-polarizing and polarizing beam splitters (BS, PBS) for spatial-mode coupling and polarization routing.
- Dove prisms (DP) for controlled spatial-mode rotations.

All components are mounted on motorized rotation stages to enable automated angle sweeps for calibration and tomography. The detection chain uses a PBS followed by single-photon detectors or power meters depending on source brightness; intensities are converted to Stokes parameters for state reconstruction.

\section{Optical Gate Realization}
Single-qubit rotations are mapped to Jones matrices realized with QWP/HWP sequences. For example, an $R_x(\pi/2)$ operation is produced with a QWP at $45^\circ$ (up to a global phase). Two-qubit interactions are emulated by coupling polarization and spatial degrees via PBS+BS networks: a polarization-controlled mode flip implements a classical-optical analogue of a CNOT between polarization and path.

The optical sequence for the two-qubit UCCSD-inspired ansatz used here is:
\begin{enumerate}
\item Prepare reference state $\ket{10}$ by applying a polarization flip on the appropriate spatial mode.
\item Apply parameterized single-qubit rotations on polarization and spatial-mode encodings using wave-plate angles $\{	heta_i\}$.
\item Entangle via PBS/BS stage that routes polarization-dependent amplitudes between spatial modes.
\item Apply a phase gate on polarization implemented as a sequence of QWPs and HWPs (Jones-phase control).
\end{enumerate}

\section{Calibration and Tomography}
Calibration proceeds in three steps: optical axis alignment, wave-plate retardance verification, and BS splitting-ratio measurement. We reconstruct polarization states by measuring Stokes parameters $I,Q,U,V$ using HWP/QWP angle settings and a PBS; each variational circuit evaluation yields the intensities needed to compute the Hamiltonian expectation value.

Tomographic reconstruction is performed assuming near-pure states and using linear inversion on the measured Stokes vector. Typical state-preparation fidelities (polarization tomography after the ansatz) are in the range 0.95--0.98 under stable alignment.

\section{Results}
We applied the optical ansatz to the two-qubit H$_2$ benchmark and measured energy estimates across the variational parameter sweep. The measured energy surface qualitatively matches the theoretical curve after applying a small calibration-phase offset. Key experimental observations:
- Repeated measurements show shot-to-shot intensity fluctuation dominated by source and coupling drift; averaging over 50 shots reduces noise to acceptable levels for optimization.
- The dominant systematic error is imperfect beam-splitter balance and residual birefringence in optics, which we partially compensate via calibration angles.

\section{Discussion}
The optical VQE prototype highlights practical trade-offs: components are inexpensive and reconfigurable, enabling rapid iterations, but analog imperfections (birefringence, alignment drift) constrain absolute accuracy. Scaling beyond a few qubits will require mode multiplexing, active stabilization, and integrated-photonics implementations to reduce footprint and improve stability.

\section{Conclusions}
We demonstrated a compact, experimentally focused translation of a small VQE circuit to a classical-optical platform. The setup provides a practical environment for testing ansatz design, calibration routines, and measurement strategies relevant to variational quantum algorithms.

\section*{Acknowledgements}
We thank the Stevens optics lab for access to equipment and helpful discussions.

\clearpage
\begin{thebibliography}{99}
\bibitem{omalley2016scalable} O'Malley, P. J. J., et al. (2016). Scalable quantum simulation of molecular energies. \textit{Physical Review X}, 6(3), 031007.
\bibitem{kandala2017hardware} Kandala, A., et al. (2017). Hardware-efficient variational quantum eigensolver for small molecules and quantum magnets. \textit{Nature}, 549(7671), 242-246.
\bibitem{jones1941new} Jones, R. C. (1941). A new calculus for the treatment of optical systems. \textit{Journal of the Optical Society of America}, 31(7), 488-493.
\end{thebibliography}

\end{document}